### 2015-09-14 article.tex
notations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper]{llncs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[utf8x]{inputenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{llncsdoc}
\usepackage[english]{babel}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{algorithmic}
\usepackage{float}
\usepackage{todonotes}
\usepackage{subfigure}
\usepackage{enumitem}
\usepackage{shuffle}

\usepackage{tikz}
\usetikzlibrary{matrix, calc, arrows, shapes.callouts}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{complexity}
\input{graph}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \renewcommand{\floatpagefraction}{.9}
% \renewcommand{\textfraction}{.1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\DeclareMathOperator{\LEFT}{\ell}
\DeclareMathOperator{\RIGHT}{r}
\DeclareMathOperator{\QQ}{\mathbb{Q}}
\DeclareMathOperator{\STD}{\mathrm{s}}
\DeclareMathOperator{\SHUFFLE}{\bullet}
\DeclareMathOperator{\BINTOPERM}{\mathrm{btp}}
\DeclareMathOperator{\PERMTOBIN}{\mathrm{ptb}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newtheorem{fact}{Fact}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\makeatletter
\newcommand{\pushright}[1]{\ifmeasuring@#1\else\omit\hfill$\displaystyle#1$\fi\ignorespaces}
\newcommand{\pushleft}[1]{\ifmeasuring@#1\else\omit$\displaystyle#1$\hfill\fi\ignorespaces}
\makeatother

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{%
Unshuffling Permutations}%

\author{%
  Samuele Giraudo \and
  St\'ephane Vialette
}% end author
\institute{%
  Universit\'e Paris-Est, LIGM (UMR 8049), CNRS, UPEM, ESIEE Paris, ENPC,
  F-77454, Marne-la-Vallée, France\\
  \email{samuele.giraudo@univ-mlv.fr} \\
  \email{vialette@univ-mlv.fr}
}% end institute
\date{\today}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
bla bla bla \ldots
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%%
%% ---- Introduction ----
%%
\section{Introduction}
\label{section:Introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% ---- Notations ----
%%
\section{Notations}
\label{section:Notations}

We follow the usual terminology on words~\cite{ChoffrutKarhumaki1997}.
Let us recall here the most important ones. The {\em empty word} is
denoted by $\epsilon$. If $u$ is a word, we denote by $\widetilde{u}$
the {\em mirror image} of $u$. If $P$ is a subset of $[|u|]$, $u_{|P}$
is the subword of $u$ consisting in the letters of $u$ at the positions
specified by the elements of $P$. The {\em shuffle} of two words $u$ and
$v$ is the set recursively defined by
$u \shuffle \epsilon = u = \epsilon \shuffle u$ and
$ua \shuffle vb = (u \shuffle vb)a \cup (ua \shuffle v)b$, were $a$ and
$b$ are letters. A word $u$ is a {\em square} if there exists a word $v$
such that $u$ belongs to $v \shuffle v$.

We denote by $S_n$ the set of permutations of size $n$ and by $S$ the
set of all permutations. In this paper, permutations are specified by
words without repetition of length $n$ on the alphabet $[n]$, so that
all above definitions about words remain valid on permutations. If
$\pi$ is a permutation, we denote by $\bar \pi$ (resp.~$\pi^{-1}$) the
{\em complementary} (resp. {\em inverse}) of $\pi$. If $u$ is a word of
integers without multiple occurrences of a same letter, $\STD(u)$ is the
{\em standarized} of $u$, that is the unique permutation of the same
size as $u$ such that for all $i, j \in [|u|]$, $u_i < u_j$ if and only
if $\STD(u)_i < \STD(u)_j$. In particular, the image of the map $\STD$
is the set $S$ of all permutations. Two words $u$ and $v$ having the
same standarized are {\em order isomorphic}. If $\sigma$ is a permutation,
there is an {\em occurrence} of (the {\em pattern}) $\sigma$ in $\pi$ if
there is a set $P$ of indexes of letters of $\pi$ such that $\sigma$ and
$\pi_{|P}$ are order isomorphic. When $\pi$ does not admit any occurrence
of $\sigma$, $\pi$ {\em avoids} $\sigma$.

For a graph $G$, we denote $\VS{G}$ as the set of vertices and
$\ES{G}$ as the set of edges.
Let $u = u_1 \, u_2 \,\ldots\, u_n$ be a word.
The \emph{clique graph associated to $u$}, denoted $K_u$,
is the complete graph on $\VS{K_u} = \{u_1, u_2, \ldots, u_n\}$.
The structure of this underlying graph is linear,
\emph{i.e.}, the set of vertices is equipped with a natural total order
$<$ defined by $u_i < u_j$ if and only if $i < j$.
(For a linear graph, we write $(u_i, u_j)$ instead of $\{u_i, u_j\}$
for an edge if it is clear from the context that $i < j$.)
Recall that two edges of a graph are \emph{independent} if they do not share
a common vertex, and that a \emph{matching} $\mathcal{M}$ in $G$ is a set of
pairwise independent edges.
A matching is \emph{perfect} if it covers all the vertices of the graph.
In case the set of vertices is equipped with a total order,
a matching $\mathcal{M}$ is said to be \emph{containment-free}
if there do not exist (independent) edges $(u_i, u_j)$ and
$(u_k, u_\ell)$ in $\mathcal{M}$ such that
$u_i < u_k < u_\ell < u_j$ or  $u_k < u_i < u_j < u_\ell$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% ---- Shuffle product on permutations ----
%%
\section{Shuffle product on permutations}
\label{section:Shuffle product on permutations}

The purpose of this section is to define a shuffle product $\SHUFFLE$
on permutations. To define this shuffle product, we shall first define
a coproduct denoted by $\Delta$, enabling to unshuffle permutations.
By duality, $\Delta$ implies the definition of $\SHUFFLE$. The reason
why we need to pass by the definition of $\Delta$ to define $\SHUFFLE$
is justified by the fact that a lot of properties of $\SHUFFLE$ depend
of properties of $\Delta$ and that this strategy allows to write concise
and clear proofs of them. We invite the reader unfamiliar with the
concepts of coproduct and duality to consult~\cite{Joni:Rota:1979}
or~\cite{Grinberg:Reiner:2014}.
\medskip

Let use denote by $\QQ[S]$ the linear span of all permutations. As
announced, we shall define on this space a kind of shuffle product and
for this purpose, we first define a linear coproduct $\Delta$ on $\QQ[S]$.
For any permutation $\pi$, we set
\begin{equation} \label{equ:unshuffling_coproduct}
    \Delta(\pi) =
    \sum_{P_1 \uplus P_2 = [|\pi|]}
    \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right).
\end{equation}
We call $\Delta$ the {\em unshuffling coproduct of permutations}. For
instance,
\begin{equation}
    \Delta(213) =
    \epsilon \otimes 213 + 2. 1 \otimes 12 +
    1 \otimes 21 + 2. 12 \otimes 1 + 21 \otimes 1 + 213 \otimes \epsilon,
\end{equation}
\begin{equation}
    \Delta(1234) =
    \epsilon \otimes 1234 + 4. 1 \otimes 123 + 6. 12 \otimes 12 +
    4. 123 \otimes 1 + 1234 \otimes \epsilon,
\end{equation}
\begin{equation}\begin{split} \label{equ:example_unshuffling_coproduct}
    \Delta(1432) & =
    \epsilon \otimes 1432 + 3. {\bf 1 \otimes 132} + 1 \otimes 321 +
    3. 12 \otimes 21 \\ & + 3. 21 \otimes 12 + 3. 132 \otimes 1 +
    321 \otimes 1 + 1432 \otimes \epsilon.
\end{split}\end{equation}
Observe that the coefficient of the tensor $1 \otimes 132$ is $3$
in~\eqref{equ:example_unshuffling_coproduct} because there are exactly
three ways to extract two disjoint subwords of the permutation $1432$
so that the letters of these two subwords have respectively the same
relative order as the letters of the permutations $1$ and $132$.
\medskip

As any coproduct, $\Delta$ leads to the definition of a product obtained
by duality in the following way. From~\eqref{equ:unshuffling_coproduct},
for any permutation $\pi$, we have
\begin{equation} \label{equ:schematic_coproduct}
    \Delta(\pi) =
    \sum_{\sigma, \nu \in S} \lambda_{\sigma, \nu}^\pi \;
    \sigma \otimes \nu,
\end{equation}
where the $\lambda_{\sigma, \nu}^\pi$ are nonnegative integers. Now,
by definition of duality, the dual product of $\Delta$ denoted by
$\SHUFFLE$ is a linear binary product on $\QQ[S]$. It satisfies, for any
permutations
$\sigma$ and $\nu$,
\begin{equation}
    \sigma \SHUFFLE \nu =
    \sum_{\pi \in S}
    \lambda_{\sigma, \nu}^\pi \; \pi,
\end{equation}
where the coefficients $\lambda_{\sigma, \nu}^\pi$ are the ones
of~\eqref{equ:schematic_coproduct}. We call $\SHUFFLE$ the
{\em shuffle product of permutations}. For instance,
\begin{equation}\begin{split} \label{equ:example_shuffle_product}
    12 \SHUFFLE 21 & =
    1243 + 1324 + 2.1342 + 2.1423 + 3. {\bf 1432} + 2134 + 2.2314 \\
    & + 3.2341 + 2413 + 2.2431 + 2.3124 + 3142 + 3.3214 + 2.3241 \\
    & + 3421 + 3.4123 + 2.4132 + 2.4213 + 4231 + 4312.
\end{split}\end{equation}
Observe that the coefficient $3$ of the permutation $1432$
in~\eqref{equ:example_shuffle_product} comes from the fact that the
coefficient of the tensor $12 \otimes 21$ is $3$
in~\eqref{equ:example_unshuffling_coproduct}.
\medskip

Intuitively, this product shuffles the values and the positions of the
letters of the permutations. There are other shuffle products on
permutations~\cite{DHT:IJAC:2002}: the {\em shifted shuffle} and the
{\em convolution product}. Our shuffle product is different from these
last two.
\medskip

We say that a permutation $\pi$ {\em appears} in the shuffle
$\sigma \SHUFFLE \nu$ of two permutations $\sigma$ and $\nu$ if the
coefficient $\lambda_{\sigma, \nu}^\pi$ defined above is different from
zero. In a more combinatorial way, this is equivalent to say that there
are two sets $P_1$ and $P_2$ of disjoints indexes of letters of $\pi$
satisfying $P_1 \sqcup P_2 = [|\pi|]$ such that the subword $\pi_{|P_1}$
is order isomorphic to $\sigma$ and the subword $\pi_{|P_2}$ is order
isomorphic to $\nu$.
\medskip

Besides, a permutation $\pi$ is a {\em square} if there is a
permutation $\sigma$ such that $\pi$ appears in $\sigma \SHUFFLE \sigma$.
In this case, we say that $\sigma$ is a {\em square root} of $\pi$.
Equivalently, $\pi$ is a square with $\sigma$ as square root if and only
if in the expansion of $\Delta(\pi)$, there is a tensor
$\sigma \otimes \sigma$ with a nonzero coefficient. In a more
combinatorial way, this is equivalent to say that there are two sets
$P_1$ and $P_2$ of disjoints indexes of letters of $\pi$ satisfying
$P_1 \sqcup P_2 = [|\pi|]$ such that the subword $\pi_{|P_1}$ and
$\pi_{|P_2}$ are order isomorphic. Computer experiments give us the
first numbers of square permutations with respects to their size, which
are, from size $0$ to $10$,
\begin{equation}
    1, 0, 2, 0, 20, 0, 504, 0, 21032, 0, 1293418.
\end{equation}
For instance, the square permutations of sizes $0$ to $4$ are
\begin{multline}
    \epsilon, 12, 21,
    1234, 1243, 1423, 1324, 1342, 4132, 3124, 3142, \\ 3412, 4312,
    2134, 2143, 2413, 4213, 2314, 2431, 4231, 3241, 3421, 4321.
\end{multline}
\medskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% ---- Binary square words and permutations ----
%%
\section{Binary square words and permutations}
\label{section:Binary square words and permutations}
In this section, we shall show that the square binary words are in
one-to-one correspondence with square permutations avoiding some
patterns. This property establishes a link between the shuffle of binary
words and our shuffle of permutations and allows to obtain a new
description of square binary words.
\medskip

Let $u$ be a binary word of length $n$ with $k$ occurrences of $0$.
We denote by $\BINTOPERM$ the map sending any such word $u$ to the
permutation obtained by replacing from left to right each occurrence of
$0$ in $u$ by $1$, $2$, \dots, $k$, and from right to left each
occurrence of $1$ in $u$ by $k + 1$, $k + 2$, \dots, $n$. For instance,
\begin{equation}
    \BINTOPERM({\bf 1}00{\bf 1}0{\bf 1}{\bf 1}0{\bf 1}000) =
    {\bf C} 12 {\bf B} 3 {\bf A} {\bf 9} 4 {\bf 8} 567.
\end{equation}
Observe that for any nonempty permutation $\pi$ in the image of
$\BINTOPERM$, there is exactly one binary word $u$ such that
$\BINTOPERM(u0) = \BINTOPERM(u1) = \pi$. In support of this observation,
when $\pi$ has an even size, we denote by $\PERMTOBIN(\pi)$ the word $ua$
such that $|ua|_0$ and $|ua|_1$ are both even, where $a \in \{0, 1\}$.
\medskip

\begin{proposition} \label{prop:bijection_binary_to_permutations_squares}
    For any $n \geq 0$, the map $\BINTOPERM$ restricted to the set of
    square binary words of length $2n$ is a bijection between this last
    set and the set of square permutations of size $2n$ avoiding the
    patterns $213$ and $231$.
\end{proposition}
\begin{proof}
    [of Proposition~\ref{prop:bijection_binary_to_permutations_squares}]
    The statement of the proposition is a consequence of the
    following claims implying that $\PERMTOBIN$ is the inverse
    map of $\BINTOPERM$ over the set of square binary words.
    \begin{claim} \label{claim:binary_to_permutation_avoiding}
        The image of $\BINTOPERM$ is the set of all permutations
        avoiding $213$ and $231$.
    \end{claim}
    \begin{proof}[of Claim~\ref{claim:binary_to_permutation_avoiding}]
        Let us first show that the image of $\BINTOPERM$ contains only
        permutations avoiding $213$ and $231$. Let $u$ be a binary word,
        $\pi = \BINTOPERM(u)$, and $P_0$ (resp. $P_1$) be the set of the
        positions of the occurrences of $0$ (resp. $1$) in $u$. By
        definition of $\BINTOPERM$, from left to right, the subword
        $v = \pi_{|P_0}$ is increasing and the subword $w = \pi_{|P_1}$
        is decreasing, and all letters of $w$ are greater than those
        of~$v$. Now, assume that $\pi$ admits an occurrence of $213$.
        Then, since $v$ is increasing and $w$ is decreasing, there is an
        occurrence of $3$ (resp. $13$, $23$) in $v$ and a relative
        occurrence of $21$ (resp. $2$, $1$). All these three cases
        contradict the fact that all letters of $w$ are greater than
        those of $v$. A similar argument shows that $\pi$ avoids~$231$
        as well.
        \smallskip

        Finally, observe that any permutation $\pi$ avoiding $213$ and
        $231$ necessarily starts by the smallest possible letter or the
        greatest possible letter. This property is then true for the
        suffix of $\pi$ obtained by deleting its first letter,
        and so on for all of its suffixes. Thus, by
        replacing each letter $a$ of $\pi$ by $0$ (resp. $1$) if $a$ has the
        role of a smallest (resp. greatest) letter, one obtains a binary
        word $u$ such that $\BINTOPERM(u) = \pi$. Hence, all permutations
        avoiding $213$ and $231$ are in the image of $\BINTOPERM$.
        \qed
    \end{proof}
    \begin{claim} \label{claim:square_binary_to_square_permutation}
        If $u$ is a square binary word, $\BINTOPERM(u)$ is a square
        permutation.
    \end{claim}
    \begin{proof}[of Claim~\ref{claim:square_binary_to_square_permutation}]
        Since $u$ is a square binary word, there is a binary word $v$
        such that $u \in v \shuffle v$. Then, there are two disjoint
        sets $P$ and $Q$ of positions of letters of $u$ such that
        $u_{|P} = v = u_{|Q}$. Now, by definition of $\BINTOPERM$, the
        words $\BINTOPERM(u)_{|P}$ and $\BINTOPERM(u)_{|Q}$ have the
        same standarized $\rho$. Hence, and by definition of the shuffle
        product of permutations, $\BINTOPERM(u)$ appears in
        $\rho \SHUFFLE \rho$, showing that $\BINTOPERM(u)$ is a square
        permutation.
        \qed
    \end{proof}
    \begin{claim} \label{claim:square_permutation_to_square_binary}
        If $\pi$ is a square permutation avoiding $213$ and $231$,
        $\PERMTOBIN(\pi)$ is a square binary word.
    \end{claim}
    \begin{proof}[of Claim~\ref{claim:square_permutation_to_square_binary}]
        Let $\pi$ be a square permutation avoiding $213$ and $231$. By
        Claim~\ref{claim:binary_to_permutation_avoiding}, $\pi$ is in
        the image of $\BINTOPERM$ and hence, $u = \PERMTOBIN(\pi)$ is a
        well-defined binary word. Since $\pi$ is a square permutation,
        there are two disjoint sets $P_1$ and $P_2$ of indexes of letters
        of $\pi$ such that $\pi_{|P_1}$ and $\pi_{|P_2}$ are order
        isomorphic. This implies, by the definitions of $\BINTOPERM$ and
        $\PERMTOBIN$, that $u_{|P_1} = u_{|P_2}$, showing that $u$ is a
        square binary word.
        \qed
    \end{proof}
\end{proof}
\medskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% ---- Algebraic issues ----
%%
\section{Algebraic issues}
\label{section:Algebraic issues}
The aim of this section is to establish some of properties of the
shuffle product of permutations $\SHUFFLE$. It is worth to note that, as
we will see, algebraic properties of the unshuffling coproduct $\Delta$
of permutations defined in
Section~\ref{section:Shuffle product on permutations} lead to
combinatorial properties of $\SHUFFLE$.
\medskip

\begin{proposition} \label{prop:shuffle_associative_commutative}
    The shuffle product $\SHUFFLE$ of permutations is associative and
    commutative.
\end{proposition}
\begin{proof}[of Proposition~\ref{prop:shuffle_associative_commutative}]
    To prove the associativity of $\SHUFFLE$, it is convenient to show
    that its dual coproduct $\Delta$ is coassociative, that is
    \begin{equation}
        (\Delta \otimes I) \Delta = (I \otimes \Delta) \Delta,
    \end{equation}
    where $I$ denote the identity map. This strategy relies on the fact
    that a product is associative if and only if its dual coproduct is
    coassociative. For any permutation $\pi$, we have
    \begin{equation} \begin{split}
    \label{equ:shuffle_associative_commutative}
        (\Delta \otimes I) \Delta(\pi) & =
        (\Delta \otimes I)
        \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right) \\
        & =
        \sum_{P_1 \uplus P_2 = [|\pi|]}
        \Delta\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes I\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & =
        \sum_{P_1 \uplus P_2 = [|\pi|]} \;
        \sum_{Q_1 \uplus Q_2 = [|P_1|]}
        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_1}\right)
        \otimes
        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_2}\right)
        \otimes \STD\left(\pi_{|P_2}\right) \\
        & =
        \sum_{P_1 \uplus P_2 \uplus P_3 = [|\pi|]}
        \STD\left(\pi_{|P_1}\right) \otimes
        \STD\left(\pi_{|P_2}\right) \otimes
        \STD\left(\pi_{|P_3}\right).
    \end{split} \end{equation}
    An analogous computation shows that $(I \otimes \Delta) \Delta(\pi)$
    is equal to the last member
    of~\eqref{equ:shuffle_associative_commutative}, whence the
    associativity of $\SHUFFLE$.
    \smallskip

    Finally, to prove the commutativity of $\SHUFFLE$, we shall show
    that $\Delta$ is cocommutative, that is for any permutation $\pi$,
    if in the expansion of $\Delta(\pi)$ there is a tensor
    $\sigma \otimes \nu$ with a coefficient $\lambda$, there is in the
    same expansion the tensor $\nu \otimes \sigma$ with the same
    coefficient $\lambda$. Clearly, a product is commutative if and only
    if its dual coproduct is cocommutative. Now, from the
    definition~\eqref{equ:unshuffling_coproduct} of $\Delta$, one
    observes that if the pair $(P_1, P_2)$ of subsets of $[|\pi|]$
    contributes to the coefficient of
    $\STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right)$,
    the pair $(P_2, P_1)$ contributes to the coefficient of
    $\STD\left(\pi_{|P_2}\right) \otimes \STD\left(\pi_{|P_1}\right)$.
    This shows that $\Delta$ is cocommutative and hence, that $\SHUFFLE$
    is commutative.
    \qed
\end{proof}
\medskip

Proposition~\ref{prop:shuffle_associative_commutative} implies that
$\QQ[S]$ endowed with the unshuffling coproduct $\Delta$ is a
coassociative cocommutative coalgebra, or in an equivalent way, that
$\QQ[S]$ endowed with the shuffle product $\SHUFFLE$ is an associative
commutative algebra.
\medskip

\begin{proposition} \label{prop:endomorphisms}
    The three linear maps
    \begin{equation}
        \phi_1, \phi_2, \phi_3 : \QQ[S] \to \QQ[S]
    \end{equation}
    linearly sending a permutation $\pi$ to, respectively,
    $\widetilde{\pi}$, $\bar \pi$, and $\pi^{-1}$ are endomorphims of
    associative algebras.
\end{proposition}
\begin{proof}[of Proposition~\ref{prop:endomorphisms}]
    To prove, for $j = 1, 2, 3$, that $\phi_j$ is a morphism of
    associative algebras, we have to prove that for all permutations
    $\sigma$ and $\nu$,
    \begin{equation}
        \phi_j(\sigma \SHUFFLE \nu) =
        \phi_j(\sigma) \SHUFFLE \phi_j(\nu).
    \end{equation}
    We shall proceed here as we have done during the proof of
    Proposition~\ref{prop:shuffle_associative_commutative} by showing
    that $\phi_j$ is a morphism of coalgebras, that is
    \begin{equation}
        \Delta \phi_j = (\phi_j \otimes \phi_j) \Delta.
    \end{equation}
    In the sequel, $\pi$ is a permutation.
    \smallskip

    If $P$ is a set of indexes of letters of $\pi$, we denote by
    $\widetilde{P}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now, since
    the operation $\widetilde{\,}$ define a bijection on the set of the
    subsets of $[|\pi|]$, and since the standardization operation
    commutes with the mirror operation on words, we have
    \begin{equation} \begin{split}
        \Delta(\phi_1(\pi))
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\phi_1(\pi)_{|P_1}\right)
        \otimes \STD\left(\phi_1(\pi)_{|P_2}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\widetilde{\pi}_{|P_1}\right)
        \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
        \otimes \STD\left(\widetilde{\pi}_{|\widetilde{P_2}}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \widetilde{\STD\left(\pi_{|P_1}\right)}
        \otimes \widetilde{\STD\left(\pi_{|P_2}\right)} \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes \phi_1\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & = (\phi_1 \otimes \phi_1) \Delta(\pi).
    \end{split} \end{equation}
    This shows that $\phi_1$ is a morphism of coalgebras and hence, that
    $\phi_1$ is a morphism of associative algebras.
    \smallskip

    Next, since by definition of the complementation operation on
    permutations, for any permutation $\tau$ and any indexes $i$ and $k$,
    we have $\tau_i < \tau_k$ if and only if $\bar \tau_i > \bar \tau_k$,
    we have
    \begin{equation} \begin{split}
        \Delta(\phi_2(\pi))
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\phi_2(\pi)_{|P_1}\right)
        \otimes \STD\left(\phi_2(\pi)_{|P_2}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\bar \pi_{|P_1}\right)
        \otimes \STD\left(\bar \pi_{|P_2}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \phi_2\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes \phi_2\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & = (\phi_2 \otimes \phi_2) \Delta(\pi).
    \end{split} \end{equation}
    This shows that $\phi_2$ is a morphism of coalgebras and hence, that
    $\phi_2$ is a morphism of associative algebras.
    \smallskip

    Finally, for any permutation $\tau$, if $P$ is a set of indexes of
    letters of $\tau$, we denote by $P_\tau^{-1}$ the set
    $\{\tau_i : i \in P\}$. Since the map sending a subset $P$ of
    $[|\pi|]$ to $P_\pi^{-1}$ is a bijection, and since
    $\STD\left(\pi_{|P}\right)^{-1} = \STD\left(\pi^{-1}_{|P_\pi^{-1}}\right)$,
    we have
    \begin{equation} \begin{split}
        \Delta(\phi_3(\pi))
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\phi_3(\pi)_{|P_1}\right)
        \otimes \STD\left(\phi_3(\pi)_{|P_2}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\pi^{-1}_{|P_1}\right)
        \otimes \STD\left(\pi^{-1}_{|P_2}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\pi^{-1}_{|{P_1}_\pi^{-1}}\right)
        \otimes \STD\left(\pi^{-1}_{|{P_2}_\pi^{-1}}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\pi_{|P_1}\right)^{-1}
        \otimes \STD\left(\pi_{|P_2}\right)^{-1} \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \phi_3\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes \phi_3\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & = (\phi_3 \otimes \phi_3) \Delta(\pi).
    \end{split} \end{equation}
    This shows that $\phi_3$ is a morphism of coalgebras and hence, that
    $\phi_3$ is a morphism of associative algebras.
    \qed
\end{proof}
\medskip

We now use the algebraic properties of $\SHUFFLE$ exhibited by
Proposition~\ref{prop:endomorphisms} to obtain combinatorial properties
of square permutations.
\medskip

\begin{proposition} \label{prop:square_stability}
    Let $\pi$ be a square permutation and $\rho$ be a square root of
    $\pi$. Then,
    \begin{enumerate}[label={\it (\roman*)},fullwidth]
        \item \label{item:square_stability_1}
        the permutation $\widetilde{\pi}$ is a square and
        $\widetilde{\rho}$ is one of its square roots;
        \item \label{item:square_stability_2}
        the permutation $\bar \pi$ is a square and $\bar \rho$ is one of
        its square roots;
        \item \label{item:square_stability_3}
        the permutation $\pi^{-1}$ is a square and $\rho^{-1}$ is one of
        its square roots.
    \end{enumerate}
\end{proposition}
\begin{proof}[of Proposition~\ref{prop:square_stability}]
    All statements~\ref{item:square_stability_1},
    \ref{item:square_stability_2}, and~\ref{item:square_stability_3} are
    direct consequences of Proposition~\ref{prop:endomorphisms}. Indeed,
    since $\pi$ is a square permutation and $\rho$ is a square root of
    $\pi$, by definition, $\pi$ appears in the product
    $\rho \SHUFFLE \rho$. Now, by Proposition~\ref{prop:endomorphisms},
    for any $j = 1, 2, 3$, since $\phi_j$ is a morphism of associative
    algebras from $\QQ[S]$ to $\QQ[S]$, $\phi_j$ commutes with the
    shuffle product of permutations $\SHUFFLE$. Hence, in particular,
    one has
    \begin{equation}
        \phi_j(\rho \SHUFFLE \rho) =
        \phi_j(\rho) \SHUFFLE \phi_j(\rho).
    \end{equation}
    Then, since $\pi$ appears in $\rho \SHUFFLE \rho$, $\phi_j(\pi)$
    appears in $\phi_j(\rho \SHUFFLE \rho)$ and also in
    $\phi_j(\rho) \SHUFFLE \phi_j(\rho)$. This shows that
    $\phi_j(\rho)$ is a square root of $\phi_j(\pi)$ and
    implies~\ref{item:square_stability_1}, \ref{item:square_stability_2},
    and~\ref{item:square_stability_3}.
    \qed
\end{proof}
\medskip

Let us make an observation about Wilf-equivalence classes of
permutations restrained on square permutations. The fact that there is a
single Wilf-equivalence class of permutations of size $3$ together with
Proposition~\ref{prop:square_stability} imply that $123$ and $321$ are
in the same Wilf-equivalence class of square permutations, and that
$132$, $213$, $231$, and $312$ are in the same Wilf-equivalence class of
square permutations. Computer experiments show us that there are two
Wilf-equivalence classes of square permutations of size $3$. Indeed, the
number of square permutations avoiding $123$ begins by
\begin{equation}
    1, 0, 2, 0, 12, 0, 118, 0, 1218, 0, 14272,
\end{equation}
while the number of square permutations avoiding $132$ begins by
\begin{equation}
    1, 0, 2, 0, 11, 0, 84, 0, 743, 0, 7108.
\end{equation}
\medskip

Besides, an other consequence of Proposition~\ref{prop:square_stability}
is that its makes sense to enumerate the sets of square permutations
quotiented by the operations of mirror image, complementary, and
inverse. The sequence enumerating these sets begins by
\begin{equation}
    1, 0, 1, 0, 6, 0, 81, 0, 2774, 0, 162945.
\end{equation}
\medskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% ---- Algorithmic issues ----
%%
\section{Algorithmic issues}
\label{section:Algorithmic issues}

\begin{proposition}
  \label{proposition:matching}
  Let $\pi \in S_{2n}$.
  Then,
  $\pi$ is the union of two order-isomorphic patterns
  if and only if
  there exists a perfect containment-free mathing
  $\mathcal{M}$ of $G_\pi$ such that
  $\LEFT(\mathcal{M}) \simeq \RIGHT(\mathcal{M})$.
\end{proposition}

The following easy lemma will prove extremely useful for simplifying the proof of
Proposition~\ref{proposition:hardness}.

\begin{lemma}
  \label{lemma:at most one edge monotone}
  Let $\pi = \pi_1 \, \sigma_1 \, \pi_2 \, \sigma_2 \, \pi_3$
  be a permutation with $n_1 = |\sigma_1| \geq 2$ and $n_2 = |\sigma_2| \geq 2$,
  and $\mathcal{M}$ be a containment-free perfect matching of $G_\pi$.
  The following assertions hold.
  \begin{enumerate}
    \item
    If $\sigma_1$ is increasing, $\sigma_2$ is decreasing and
    $\sigma_1$ is above $\sigma_2$
    (see Fig.~\ref{subfig:increasing before and aboce decreasing}),
    then $\mathcal{M}$ sontains at most one $(\sigma_1, \sigma_2)$-edge.
    \item
    If $\sigma_1$ is decreasing, $\sigma_2$ is increasing and
    $\sigma_1$ is below $\sigma_2$
    (see Fig.~\ref{subfig:decresing before and below increasing}),
    then $\mathcal{M}$ sontains at most one $(\sigma_1, \sigma_2)$-edge.
  \end{enumerate}
\end{lemma}

\begin{proof}[of Lemma~\ref{lemma:at most one edge monotone}]
  \begin{figure}[t!]
    \centering
    \subfigure[%
      An increasing pattern before and above a decreasing pattern.
    ]{%
      \begin{tikzpicture}
        [
          scale=0.45,
          label/.style={anchor=base}
        ]
        % increasing pattern
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,5.8) grid (5.2,11.2);
        \foreach \x/\y in {0/6,1/7,2/8,3/9,4/10,5/11} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % decreasing pattern
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
        \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % labels
        \node [label] (a) at (1,5) {$a$};
        \node [label] (b) at (4,5) {$b$};
        \node [label] (ap) at (7,-1) {$a'$};
        \node [label] (bp) at (10,-1) {$b'$};
        \node (nu1) at (2.5,4) {$\sigma_1$};
        \node (nu2) at (8.5,-2) {$\sigma_2$};
        % crossings edges
        \draw [line width=1pt]
        (1,7) .. controls +(0,12) and +(0,10) .. (7,4);
        \draw [line width=1pt]
        (4,10) .. controls +(0,12) and +(0,10) .. (10,1);
      \end{tikzpicture}
      \label{subfig:increasing before and aboce decreasing}
    }% end subfigure
    \qquad
    \subfigure[%
      A decreasing pattern before and below an increasing pattern.
    ]{%
    \begin{tikzpicture}
      [
        scale=0.45,
        label/.style={anchor=base}
      ]
      % decreasing pattern
      \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (5.2,5.2);
      \foreach \x/\y in {0/5,1/4,2/3,3/2,4/1,5/0} {
          \draw [fill=black] (\x,\y) circle (0.1);
      }
      % decreasing pattern
      \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,5.8) grid (11.2,11.2);
      \foreach \x/\y in {6/6,7/7,8/8,9/9,10/10,11/11} {
          \draw [fill=black] (\x,\y) circle (0.1);
      }
      % labels
      \node [label] (a) at (1,-1) {$a$};
      \node [label] (b) at (4,-1) {$b$};
      \node [label] (ap) at (7,5) {$a'$};
      \node [label] (bp) at (10,5) {$b'$};
      \node (nu1) at (2.5,-2) {$\sigma_1$};
      \node (nu2) at (8.5,4) {$\sigma_2$};
      % crossings edges
      \draw [line width=1pt]
      (1,4) .. controls +(0,10) and +(0,12) .. (7,7);
      \draw [line width=1pt]
      (4,1) .. controls +(0,10) and +(0,12) .. (10,10);
      \end{tikzpicture}
      \label{subfig:decresing before and below increasing}
    }% end subfigure
    \caption{\label{fig:subfig:no (nu_1, nu_2)-edge}%
      Lemma~\ref{lemma:at most one edge monotone}:
      Assuming a $(\nu_1,\nu_2)$-edge $(a, b)$ in $\mathcal{M}$.
    }
  \end{figure}
  We first prove \emph{1}.
  Suppose, aiming at a contradiction, that there exist two
  $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in $\mathcal{M}$.
  Since $\mathcal{M}$ is containment-free, we may assume $a < b$ and $a' < b'$.
  Therefore, we have
  $\pi(a) < \pi(b)$ (since $\sigma_1$ is increasing) and
  $\pi(a') > \pi(b')$ (since $\sigma_2$ is decreasing),
  which contradicts our hypothesis about $\mathcal{M}$.

  As for \emph{2}, suppose, still aiming at a contradiction,
  that there exist two
  $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in $\mathcal{M}$.
  Since $\mathcal{M}$ is containment-free, we may assume $a < b$ and $a' < b'$.
  Therefore, we have
  $\pi(a) > \pi(b)$ (since $\sigma_1$ is decreasing) and
  $\pi(a') < \pi(b')$ (since $\sigma_2$ is increasing),
  which contradicts our hypothesis about $\mathcal{M}$.
  \qed
\end{proof}

\begin{proposition}
  \label{proposition:hardness}
  Deciding whether a permutation is the union of two
  order-isomorphic patterns is \NPC.
\end{proposition}

\begin{proof}
  For the sake of clarity, we will that a bunch of positions $P$
  is \emph{above} (resp. \emph{below}) above another bunch of positions $P'$
  in permutation $\mu$
  if $\mu(x) > \mu(y)$ (resp. $\mu(x) < \mu(y)$) for every $x \in P$
  and every $y \in P'$.
  For exemple, $\nu_1$ is above $\nu_2$ but below $\nu'_1$ in permutation $\mu$.

  \begin{figure}[t!]
    \centering
    \begin{tikzpicture}[
      scale=.95,
      >=stealth',
      shorten >=1pt,
      main node/.style={align=center},
      cell/.style={draw,ultra thick,fill=black!5},
      structure link/.style={line width=1.5pt},
      pattern link/.style={line width=1.5pt,black!20},
      monotone/.style={->,thick}
      ]
      % edges
      % pi 1 - pi 2
      \draw [pattern link] (8.5,6) .. controls +(0,1) and +(0,2) .. (10.5,5);
      % pi 2 - sigma 2
      \draw [pattern link] (10.5,5) .. controls +(0,1) and +(0,3) .. (11.5,4);
      % 1
      \draw [structure link] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
      \node (claim edge 1) at (1.5,13.5) {Claim~\ref{claim:(nu_1, nu'_1)-edge}};
      \node (claim edge 1) at (1.5,13) {Claim~\ref{claim:all (nu_1, nu'_1)-edges}};
      % 2
      \draw [structure link] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
      \node (claim edge 2) at (4,4) {Claim\ref{claim:(nu_2, nu'_2)-edge}};
      \node (claim edge 2) at (4,3.5) {Claim\ref{claim:all (nu_2, nu'_2)-edges}};
      % 3
      \draw [structure link] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
      \node (claim edge 1) at (5.5,11.5) {Claim\ref{claim:(nu_3, nu'_3)-edge}};
      % 4
      \draw [structure link] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
      \node (claim edge 1) at (8.5,8.5) {Claim\ref{}};
      % sigma 1 - pi 1
      \draw [structure link] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
      \node (claim edge 1) at (6.5,4) {Claim\ref{}};
      % nodes
      % 1
      \draw [cell] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
      \draw [monotone] (0,10) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_1$};
      %\draw [draw,ultra thick,fill=black!5] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
      \draw [cell] (2,11) -- (3,11) -- (3,12) -- (2,12) -- cycle;
      \draw [monotone] (2,11) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_1$};
      % 2
      \draw [cell] (1,1) -- (2,1) -- (2,2) -- (1,2) -- cycle;
      \draw [monotone] (1,2) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_2$};
      \draw [cell] (6,0) -- (7,0) -- (7,1) -- (6,1) -- cycle;
      \draw [monotone] (6,1) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_2$};
      % 3
      \draw [cell] (3,8) -- (4,8) -- (4,9) -- (3,9) -- cycle;
      \draw [monotone] (3,8) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_3$};
      \draw [cell] (7,9) -- (8,9) -- (8,10) -- (7,10) -- cycle;
      \draw [monotone] (7,9) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_3$};
      % 4
      \draw [cell] (5,7) -- (6,7) -- (6,8) -- (5,8) -- cycle;
      \draw [monotone] (5,8) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_4$};
      \draw [cell] (9,2) -- (10,2) -- (10,3) -- (9,3) -- cycle;
      \draw [monotone] (9,3) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_4$};
      % sigma 1
      \draw [cell] (4,6) -- (5,6) -- (5,7) -- (4,7) -- cycle;
      \node [main node] (sigma1) at (4.5,6.5) {$\sigma'$};
      % pi 1
      \draw [cell] (8,5) -- (9,5) -- (9,6) -- (8,6) -- cycle;
      \node [main node] (pi1) at (8.5,5.5) {$\pi'$};
      % pi 2
      \draw [cell] (10,4) -- (11,4) -- (11,5) -- (10,5) -- cycle;
      \node [main node] (sigma2) at (10.5,4.5) {$\pi''$};
      % sigma 2
      \draw [cell] (11,3) -- (12,3) -- (12,4) -- (11,4) -- cycle;
      \node [main node] (pi2) at (11.5,3.5) {$\sigma''$};
      % grid
      \draw[step=1cm,black!30,ultra thin,fill=black!10] (0,0) grid (12,12);
      % row size
      \foreach \y/\N in {0.5/N_2,1.5/N_2,2.5/N_4,3.5/k,4.5/n,5.5/n+2,6.5/k+2,7.5/N_4,8.5/N_3,9.5/N_3,10.5/N_1,11.4/N_1} {
          \node at (-0.5,\y) (R\y) {$\N$};
      }
      % column size
      \foreach \x/\N in
      {0.5/N_1,1.5/N_2,2.5/N_1,3.5/N_3,4.5/k+2,5.5/N_4,6.5/N_2,7.5/N_3,8.5/n+2,9.5/N_4,10.5/n,11.5/k} {
          \node at (\x,-0.5) (C\x) {$\N$};
      }
    \end{tikzpicture}
    \caption{\label{fig:reduction}}
  \end{figure}

  \begin{align*}
  N_4 &= 2(2n + k + 2) + 1  = 4n + 2k + 5 \\
  N_3 &= 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 12k + 9 \\
  N_2 &= 2(2N_3 + 2N_4 + 2n + 2k + 4) + 1 \\
  N_1 &= 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1\text{.}
  \end{align*}
  Notice that $N_1, N_2, N_3$ and $N_4$ are polynomial in $n$.
  The crucial properties are that
  (i) $N_1, N_2, N_3$ and $N_4$ are odd integers
  and
  (ii) $N_i > \left(\sum_{i < j \leq 4} 2N_j\right) + 2n + 2k + k$
  for every $1 \leq i \leq 4$.

  We now turn to defining various gadgets (sequences of integers)
  that act as building blocks in our construction:
  \begin{align*}
  \sigma'  &= ((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2] \\
  \pi'     &= ((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2] \\
  \sigma'' &= \sigma \; [2N_2 + N_4] \\
  \pi''    &= \pi \; [2N_2 + N_4 + k] \\
  \nu_1    &= \nearrow_{N_1} \; [2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
  \nu'_1   &= \nearrow_{N_1} \; [N_1 + 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
  \nu_2    &= \nearrow_{N_2} \; [N_2] \\
  \nu'_2   &= \searrow_{N_2} \\
  \nu_3    &= \nearrow_{N_3} \; [2N_2 + 2N_4 + 2n + 2k + 4] \\
  \nu'_3   &= \nearrow_{N_3} \; [2N_2 + N_3 + 2N_4 + 2n + 2k + 4] \\
  \nu_4    &= \searrow_{N_4} \; [2N_2 + N_4 + 2n + 2k + 4] \\
  \nu'_4   &= \searrow_{N_4} \; [2N_2]\text{.}
  \end{align*}
  We are now in position to define our target permutation $\mu$:
  $$
  \mu
  =
  \nu_1 \; \nu_2 \; \nu'_1 \; \nu_3 \; \sigma' \; \nu_4 \; \nu'_3 \; \pi' \; \nu'_4 \; \pi'' \; \sigma''
  \text{.}
  $$
  We claim that $\sigma$ occurs as a pattern in $\pi$ if and only if
  there exists a containment-free perfect matching
  $\mathcal{M}$ in $G_\mu$ such that
  for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.

  Suppose first that $\sigma$ occurs in $\pi$ as a pattern.
  We show how to construct a containment-free perfect matching
  $\mathcal{M}$ in $G_\mu$ such that
  for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
  \begin{itemize}
    \item $\mathcal{M}$ contains $N_1$ pairwise crossing
    $(\nu_1, \nu'_1)$-edges.
    \item $\mathcal{M}$ contains $N_2$ pairwise crossing
    $(\nu_2, \nu'_2)$-edges.
    \item $\mathcal{M}$ contains $N_3$ pairwise crossing
    $(\nu_3, \nu'_3)$-edges.
    \item $\mathcal{M}$ contains $N_4$ pairwise crossing
    $(\nu_4, \nu'_4)$-edges.
    \item $\mathcal{M}$ contains $k+2$ pairwise crossing
    $(\sigma', \pi')$-edges as depicted in Figure~\ref{}.
    (Notice that we use here the fact that $\sigma$ occurs as a pattern in $\pi$).
    \item $\mathcal{M}$ contains $n-k$ pairwise crossing
    $(\pi', \pi'')$-edges as depicted in Figure~\ref{}.
    \item $\mathcal{M}$ contains $k$ pairwise crossing
    $(\pi'', \sigma'')$-edges as depicted in Figure~\ref{}.
    (Notice that we use here again the fact that $\sigma$ occurs as a pattern in $\pi$).
  \end{itemize}
  It can be easily checked (probably referring to Figure~\ref{}) that
  $\mathcal{M}$ is perfect and that
  for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.

    \begin{figure}[t!]
      \centering
        \begin{tikzpicture}
          [
            scale=0.45,
            label/.style={anchor=base},
            cell/.style={draw,thick,fill=black!5}
          ]
          % sigma'
          \draw [cell,rounded corners] (1.5,16.5) -- (4.5,16.5) -- (4.5,19.5) -- (1.5,19.5) -- cycle;
          \draw[step=1cm,black!60,ultra thin,fill=black!10] (0.8,16.8) grid (5.2,21.2);
          \foreach \x/\y in {1/20,2/19,3/17,4/18,5/21} {
              \draw [fill=black] (\x,\y) circle (0.1);
          }
          \node (pip) at (3,22) {$\sigma'$};
          \node[%
            draw,
            text width=5.5cm,text justified,color=black!50,rounded corners,
            rectangle callout,callout relative pointer={(0.0cm,-1.5cm)}
          ] at (3,26)
          {%
            $((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2]$
          };%
          %\node at (3,-0.5) {$((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2]$};
          \node at (-0.5,20) {$k+1$};
          \node at (-0.5,21) {$k+2$};
          % pi'
          \draw [cell,rounded corners] (6.5,8.5) -- (12.5,8.5) -- (12.5,14.5) -- (6.5,14.5) -- cycle;
          \draw[step=1cm,black!60,ultra thin,fill=black!10] (5.8,8.8) grid (13.2,16.2);
          \foreach \x/\y in {6/15,7/12,8/13,9/10,10/9,11/11,12/14,13/16} {
              \draw [fill=black] (\x,\y) circle (0.1);
          }
          \node (pip) at (9.5,17) {$\pi'$};
          \node[%
            draw,
            text width=5.5cm,text justified,color=black!50,rounded corners,
            rectangle callout,callout relative pointer={(0.0cm,3cm)}
          ] at (9.5,0)
          {%
            $((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2]$
          };%
          %\node at (9.5,-2) {$((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2]$};
          \node at (-0.5,15) {$n+1$};
          \node at (-0.5,16) {$n+2$};
          % pi''
          \draw [cell,rounded corners] (13.5,2.5) -- (19.5,2.5) -- (19.5,8.5) -- (13.5,8.5) -- cycle;
          \draw[step=1cm,black!60,ultra thin,fill=black!10] (13.8,2.8) grid (19.2,8.2);
          \foreach \x/\y in {14/6,15/7,16/4,17/3,18/5,19/8} {
              \draw [fill=black] (\x,\y) circle (0.1);
          }
          \node (pipp) at (16.75,9) {$\pi''$};
          \node[%
            draw,
            text width=2.25cm,text justified,color=black!50,rounded corners,
            rectangle callout,callout relative pointer={(0.0cm,-4cm)}
          ] at (16.5,19)
          {%
            $\pi [2N_2 + N_4 + k]$
          };%
          %\node at (16.5,-3.5) {$\pi [2N_2 + N_4 + k]$};
          % sigma''
          \draw [cell,rounded corners] (19.5,-0.5) -- (22.5,-.5) -- (22.5,2.5) -- (19.5,2.5) -- cycle;
          \draw[step=1cm,black!60,ultra thin,fill=black!10] (19.8,-0.2) grid (22.2,2.2);
          \foreach \x/\y in {20/2,21/0,22/1} {
              \draw [fill=black] (\x,\y) circle (0.1);
          }
          \node (sigmapp) at (21,3) {$\sigma''$};
          \node[%
            draw,
            text width=1.75cm,text justified,color=black!50,rounded corners,
            rectangle callout,callout relative pointer={(0.0cm,1.25cm)}
          ] at (21,-4.5)
          {%
            $\sigma [2N_2 + N_4]$
          };%
          %\node at (21,-5) {$\sigma [2N_2 + N_4]$};
          % edges sigma' - pi'
          \draw [line width=1pt]
          (1,20) .. controls +(0,9) and +(0,4) .. (6,15);
          \draw [line width=1pt]
          (5,21) .. controls +(0,9) and +(0,4) .. (13,16);
          \draw [line width=1pt]
          (2,19) .. controls +(0,-17) and +(0,-7) .. (8,13);
          \draw [line width=1pt]
          (3,17) .. controls +(0,-17) and +(0,-7) .. (9,10);
          \draw [line width=1pt]
          (4,18) .. controls +(0,-17) and +(0,-7) .. (11,11);
          % edges pi' - pi''
          \draw [line width=1pt]
          (7,12) .. controls +(0,9) and +(0,4) .. (14,6);
          \draw [line width=1pt]
          (10,9) .. controls +(0,9) and +(0,4) .. (17,3);
          \draw [line width=1pt]
          (12,14) .. controls +(0,9) and +(0,4) .. (19,8);
          % edges pi'' - sigma''
          \draw [line width=1pt]
          (15,7) .. controls +(0,-9) and +(0,-4) .. (20,2);
          \draw [line width=1pt]
          (16,4) .. controls +(0,-9) and +(0,-4) .. (21,0);
          \draw [line width=1pt]
          (18,5) .. controls +(0,-9) and +(0,-4) .. (22,1);
          % text
          % \node[%
          %   draw,
          %   text width=5.5cm,text justified,color=black!50,
          %   rectangle callout,callout relative pointer={(0.5cm,-.75cm)}
          % ] at (1,27)
          % {%
          % $2$ pairwise crossing $(\sigma',\pi')$-edges that
          % link the guards
          % };%
          % \node[%
          %   draw,
          %   text width=5.5cm,text justified,color=black!50,
          %   rectangle callout,callout relative pointer={(-0.1cm,-2.5cm)}
          % ] at (17,24)
          % {%
          % $n-k$ pairwise crossing $(\sigma',\pi')$-edges that
          % denote positions that do not correspond to
          % a specific occurence of $\sigma$ in $\pi$
          % (at positions $2$, $3$ and $5$).
          % };%
          % \node[%
          %   draw,
          %   text width=5.5cm,text justified,color=black!50,
          %   rectangle callout,callout relative pointer={(0.75cm,1.5cm)}
          % ] at (2.5,4,5)
          % {%
          % $k$ pairwise crossing $(\sigma',\pi')$-edges that
          % denote a specific occurence of $\sigma$ in $\pi$
          % (at positions $2$, $3$ and $5$).
          % };%
          % \node[%
          %   draw,
          %   text width=5.5cm,text justified,color=black!50,
          %   rectangle callout,callout relative pointer={(2cm,.75cm)}
          % ] at (7,-3)
          % {%
          % $k$ pairwise crossing $(\pi''; \sigma'')$-edges that
          % denote a specific occurence of $\sigma$ in $\pi$
          % (at positions $2$, $3$ and $5$).
          % };%
        \end{tikzpicture}
        \caption{\label{fig:subfig:sigma' - pi' - pi'' - sigma''}%
        $\sigma = 312$ and $\pi = 4\mathbf{5}\mathbf{2}1\mathbf{3}6$
        (where a specific occurrence of $\sigma$ in $\pi$ is depicted in bold).
        }
      \end{figure}

  More precisely,
  (i) the first integer of $\sigma'$
  (\emph{i.e.}, $(k+1) + (2N_2 + N_4 + 2n + k + 2)$) is linked
  to the first integer of $\pi'$
  ((\emph{i.e.}, $(n+1) + (2N_2 + N_4 + n + k + 2)$)),
  (ii)

  Conversely, suppose that there exists a containment-free perfect matching
  $\mathcal{M}$ in $G_\mu$ such that for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
  We show that $\sigma$ occurs as a pattern in $\pi$.
  Whereas the matching $\mathcal{M}$ may not be as regular as in the forward
  direction, the main idea is to prove that $\mathcal{M}$ contains enough structure
  (more precisely, enough $(\sigma', \pi')$-edges) so that we can conclude.
  We have divided the reverse direction into a serie of basic claims that
  progressively defines the overall structure of $\mathcal{M}$.

    \begin{figure}[t!]
      \centering
      \subfigure[%
        A $(\nu_1, \nu_2)$-edge $(a, b)$ together with a
        $(\nu_2, \nu'_1)$-edge $(x, y)$.
        Notice that we may have $x<b$
        (as long as $N_1 < x \leq N_1+N_2$ and $N_1 < b \leq N_1+N_2$
        do hold).
      ]{%
        \begin{tikzpicture}
          [
            scale=0.25,
            label/.style={anchor=base}
          ]
          % nu_1
          \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
          \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
              \draw [fill=black] (\x,\y) circle (0.1);
          }
          % nu_4
          \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
          \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
              \draw [fill=black] (\x,\y) circle (0.1);
          }
          % nu'_1
          \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,15.8) grid (17.2,21.2);
          \foreach \x/\y in {12/16,13/17,14/18,15/19,16/20,17/21} {
              \draw [fill=black] (\x,\y) circle (0.1);
          }
          % labels
          \node [label] (y) at (17,14.5) {$y$};
          \node [label] (yp) at (10,-1.5) {$y'$};
          \node [label] (a) at (0,8.5) {$a$};
          \node [label] (ap) at (8,-1.5) {$a'$};
          \node (nu1) at (2.5,7.7) {$\nu_1$};
          \node (nu2) at (8.5,-2.5) {$\nu_2$};
          \node (nu1') at (14.5,13.5) {$\nu'_1$};
          % crossings edges
          \draw [line width=1pt]
          (1,11) .. controls +(0,12) and +(0,4) .. (12,16);
          \draw [line width=1pt]
          (2,12) .. controls +(0,12) and +(0,4) .. (13,17);
          \draw [line width=1pt]
          (3,13) .. controls +(0,12) and +(0,4) .. (14,18);
          \draw [line width=1pt]
          (4,14) .. controls +(0,12) and +(0,4) .. (15,19);
          \draw [line width=1pt]
          (5,15) .. controls +(0,12) and +(0,4) .. (16,20);
          % data edges
          \draw [line width=1pt]
          (0,10) .. controls +(0,8) and +(0,20) .. (8,3);
          \draw [line width=1pt]
          (10,1) .. controls +(0,32) and +(0,10) .. (17,21);
        \end{tikzpicture}
        \label{subfig:no (nu_1, nu_2)-edge - 1}
      }% end subfigure
      \qquad
      \subfigure[%
        A $(\nu_1, \nu'_1)$-edge $(x', y')$ together with an
        edge $(y, z)$ which is neither
        a $(\nu_1, \nu'_1)$-edge,
        nor a $(\nu_2, \nu'_1)$-edge,
        nor a $(\nu'_1, \nu'_1)$-edge.
        Notice that $y$ may be any position satisfying
        $N_1+N_2 < y \leq 2N_1+N_2$.
      ]{%
      \begin{tikzpicture}
        [
          scale=0.25,
          label/.style={anchor=base}
        ]
        % nu_1
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
        \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % nu_4
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
        \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % nu'_1
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,15.8) grid (17.2,21.2);
        \foreach \x/\y in {12/16,13/17,14/18,15/19,16/20,17/21} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % labels
        \node [label] (y) at (17,14.5) {$y$};
        \node [label] (ypp) at (23,-4.5) {$y''$};
        \node [label] (a) at (0,8.5) {$a$};
        \node [label] (a) at (2,8.5) {$x'$};
        \node [label] (a) at (14,14.5) {$y'$};
        \node [label] (ap) at (8,-1.5) {$a'$};
        \node (nu1) at (2.5,7.7) {$\nu_1$};
        \node (nu2) at (8.5,-2.5) {$\nu_2$};
        \node (nu1') at (14.5,13.5) {$\nu'_1$};
        % crossings edges
        \draw [line width=1pt]
        (1,11) .. controls +(0,12) and +(0,4) .. (12,16);
        \draw [line width=1pt]
        (2,12) .. controls +(0,12) and +(0,4) .. (13,17);
        \draw [line width=1pt]
        (3,13) .. controls +(0,12) and +(0,4) .. (14,18);
        \draw [line width=1pt]
        (4,14) .. controls +(0,12) and +(0,4) .. (15,19);
        \draw [line width=1pt]
        (5,15) .. controls +(0,12) and +(0,4) .. (16,20);
        % data edges
        \draw [line width=1pt]
        (0,10) .. controls +(0,8) and +(0,20) .. (8,3);
        \draw [line width=1pt]
        (17,21) .. controls +(0,10) and +(0,35) .. (23,-3);
        \end{tikzpicture}
        \label{subfig:no (nu_1, nu_2)-edge - 2}
      }% end subfigure
      \caption{\label{fig:subfig:no (nu_1, nu_2)-edge}%
        Claim~\ref{claim:no (nu_1, nu_2)-edge}:
        Assuming a $(\nu_1,\nu_2)$-edge $(a, b)$ in $\mathcal{M}$.
      }
    \end{figure}

  \begin{claim}
    \label{claim:no (nu_1, nu_2)-edge}
    There is no $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_2)-edge}]
    First, according to Lemma~\ref{lemma:at most one edge monotone},
    there exists at most one $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
    Suppose now, aiming at a contradiction, that there exists
    exactly one $(\nu_1, \nu_2)$-edge $e = (a, a')$ in $\mathcal{M}$.
    Under this assumption, we show that $\mathcal{M}$ does not contain any $(\nu_1, \nu_1)$-edge.
    Indeed, suppose, aiming at a contradiction,
    that there exists a $(\nu_1, \nu_1)$-edge $(x, x')$ in $\mathcal{M}$.
    Since $\mathcal{M}$ is containment-free,
    we must have $x < a$.
    Then it follows that $\mu(x) < \mu(a)$ (since $\nu_1$ is increasing) and
    $\mu(x') > \mu(a')$ (since $\nu_1$ is above $\nu_2$),
    which contradicts our hypothesis about $\mathcal{M}$.
    Hence $\mathcal{M}$ does not contain any $(\nu_1, \nu_1)$-edge.
    Furthermore, since $\nu_1$ is increasing and
    there exists only one $(\nu_1, \nu_2)$-edge $(a, a')$ in $\mathcal{M}$,
    we conclude that
    $\mathcal{M}$ contains $N_1-1$ pairwise crossing $(\nu_1, \nu'_1)$-edges.
    But $|\nu_1| = |\nu'_1| = N_1$ and $\mathcal{M}$ is perfect,
    and hence there exists a position, say $y$, in $\nu'_1$ that is
    not involved in a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
    We rule out this configuration by considering two cases:
    \begin{itemize}
      \item Position $y$ is involved in a $(\nu_2, \nu'_1)$-edge,
      say $(y', y)$. Then, we have $\mu(a) > \mu(y')$ (since $\nu_1$ is above $\nu_2$)
      and $\mu(a') < \mu(y)$ (since $\nu_2$ is below $\nu'_1$),
      which contradicts our hypothesis about $\mathcal{M}$.
      See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 1}.
      \item Position $y$ is not involved in a $(\nu_2, \nu'_1)$-edge.
      Since $\mathcal{M}$ contains $N_1-1$ pairwise crossing $(\nu_1, \nu'_1)$-edges
      in $\mathcal{M}$ and $|\nu'_1| = N_1$, there exists an edge $(y, y'')$ in $\mathcal{M}$
      that is neither a $(\nu_1, \nu'_1)$-edge nor a $(\nu_2, \nu'_1)$-edge
      nor a $(\nu'_1, \nu'_1)$-edge.
      In other words, $y'' > 2N_1 + N_2$.
      Now, for any $(\nu_1, \nu'_1)$-edge $(x, x')$ in $\mathcal{M}$, we
      have $\mu(x) < \mu(y)$ (since $\nu_1$ is below $\nu'_1$)
      and $\mu(x') > \mu(y'')$ (since $\nu'_1$ is above any other gadget),
      which contradicts our hypothesis about $\mathcal{M}$.
      See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 2}.
    \end{itemize}
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:(nu_1, nu'_1)-edge}
    There is at least one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:(nu_1, nu'_1)-edge}]
    Suppose, aiming at a contradiction, that there is no
    $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
    Then it follows that there exists an edge $(x, x')$ in $\mathcal{M}$,
    $1 \leq x \leq N_1$, that is neither a
    $(\nu_1, \nu_1)$-edge (since $N_1$ is odd)
    nor a $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge})
    nor a $(\nu_1, \nu'_1)$ (by our contradiction hypothesis).
    (In other words, $x \leq N_1$ and $x' > 2N_1 + N_2$.)
    Therefore, since $\mathcal{M}$ is containment-free,
    there is neither a $(\nu_2, \nu_2)$-edge nor a $(\nu'_1, \nu'_1)$-edge
    in $\mathcal{M}$.
    Then it follows that every position $y$ of $\nu'_1$
    (\emph{i.e.}, $N_1+N_2 < y \leq 2N_1+N_2$)
    is involved in an edge $(y, y'$ of $\mathcal{M}$ with $y' \geq 2N_1+N_2$
    (otherwise $\mathcal{M}$ would not be containment-free).
    But $|\nu'_1| = N_1 = 2(2N_2 + 2n_3 + 2N_4 + 2n + 2k + 4) + 1
    > 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4$, and hence at least one
    position in $\nu'_1$ can not be involved in an edge in $\mathcal{M}$.
    Then it follows that $\mathcal{M}$ is not a perfect matching,
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  The above claim will be complement in upcoming Claim~\ref{claim: all (nu_1, nu'_1)-edges}.

  \begin{claim}
    \label{claim:no (nu_2, nu_2)-edge}
    There is no $(\nu_2, \nu_2)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu_2)-edge}]
  Combine Claim~\ref{claim:(nu_1, nu'_1)-edge} with
  the fact that $\mathcal{M}$ is containment-free.
  \qed
  \end{proof}

  \begin{claim}
    \label{claim:no (nu_2, nu'_1)-edge}
    There is no $(\nu_2, \nu'_1)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu'_1)-edge}]
    First, according to Lemma~\ref{lemma:at most one edge monotone},
    there exists at most one $(\nu_2, \nu'_1)$-edge in $\mathcal{M}$.
    Suppose now, aiming at a contradiction, that there exists
    exactly one $(\nu_2, \nu'_1)$-edge $e = (a, a')$ in $\mathcal{M}$.
    According to Claim~\ref{claim:(nu_1, nu'_1)-edge}, there exists at least
    one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
    Furtheremore, since
    there is no $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge}),
    every position in $\nu_1$ is involved either in a
    $(\nu_1, \nu_1)$-edge or in a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$
    (otherwise $\mathcal{M}$ would not be containment-free).
    Then it follows that that there is an odd number of $(\nu_1, \nu'_1)$-edges
    in $\mathcal{M}$ since $\mathcal{M}$ is perfect and $|\nu_1| = N_1$ is odd.
    Combining this with the fact that there exists exactly one $(\nu_2, \nu'_1)$-edge
    in $\mathcal{M}$ and $|\nu'_1| = N_1$ is odd,
    we conclude that there exists a position, say $x$, in $\nu'_1$
    that is neither involved in a $(\nu_1, \nu'_1)$-edge nor a $(\nu_1, \nu'_1)$-edge
    nor a $(\nu'_1, \nu'_1)$-edge.
    Write $(x, x')$, $N_1+N_2 < x \leq 2N_1 + N_2$ and $x' > 2N_1 + N_2$, this edge.
    We have $\mu(a) < \mu(x)$ (since $\nu_2$ is below $\nu'_1$)
    and $\mu(a') > \mu(x')$ (since $\nu'_1$ is above any other gadget),
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:(nu_2, nu'_2)-edge}
    There is at least one $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:(nu_2, nu'_2)-edge}]
    Suppose, aiming at a contradiction that there is no
    $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
    Notice that there is neither
    a $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge})
    nor a $(\nu_2, \nu_2)$-edge (Claim~\ref{claim:no (nu_2, nu_2)-edge})
    nor a $(\nu_2, \nu'_1)$-edge (Claim~\ref{claim:no (nu_2, nu'_1)-edge})
    in $\mathcal{M}$.
    But $|\nu_2| = N_2 > 2N_3 + 2N_4 + 2n + 2k +4$.
    Hence $\mathcal{M}$ is not a perfect matching,
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:no edge below (nu_2, nu'_2)-edge}
    There is
    neither a $(\nu'_1, \nu'_1)$-edge,
    nor a $(\nu'_1, \nu_3)$-edge,
    nor a $(\nu'_1, \sigma')$-edge,
    nor a $(\nu'_1, \nu_4)$-edge
    nor a $(\nu_3, \nu_3)$-edge,
    nor a $(\nu_3, \sigma')$-edge,
    nor a $(\nu_3, \nu_4)$-edge,
    nor a $(\sigma', \sigma')$-edge,
    nor a $(\sigma', \nu_4)$-edge,
    nor a $(\nu_4, \nu_4)$-edge
    in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge}]
    Combine Claim~\ref{claim:(nu_2, nu'_2)-edge} with the fact that
    $\mathcal{M}$ is containment-free.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim: no edge right of nu'_2}
    There is
    neither a $(\nu'_2, \nu'_3)$-edge,
    nor a $(\nu'_2, \pi')$-edge,
    nor a $(\nu'_2, \nu'_4)$-edge,
    nor a $(\nu'_2, \pi'')$-edge
    nor a $(\nu'_2, \sigma'')$-edge
    in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim: no edge right of nu'_2}]
    Suppose, aiming at a contradiction, that
    $\mathcal{M}$ contains an edge $(a, a')$ which is
    either a $(\nu'_2, \nu'_3)$-edge
    or a $(\nu'_2, \pi')$-edge
    or a $(\nu'_2, \nu'_4)$-edge
    or a $(\nu'_2, \pi'')$-edge
    or a $(\nu'_2, \sigma'')$-edge.
    According to Claim~\ref{claim:(nu_2, nu'_2)-edge},
    there exists a $(\nu_2, \nu'_2)$-edge $(b, b')$ in $\mathcal{M}$.
    Then it follows that we have
    $\mu(b) > \mu(a)$ (since $\nu_2$ is above $\nu'_2$) and
    $\mu(b') < \mu(a')$ (since $\nu'_3$, $\pi'$, $\nu'_4$, $\pi''$ and
    $\sigma''$ are all above $\nu'_2$),
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  % \begin{claim}
  %   \label{claim:no (nu_1, nu'_2)-edge}
  %   There is no $(\nu_1, \nu'_2)$-edge in $\mathcal{M}$.
  % \end{claim}
  %
  % \begin{proof}[of Claim~\ref{claim:no (nu_1, nu'_2)-edge}]
  %   Suppose, aiming at a contradiction that there is a
  %   $(\nu_1, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
  %   According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
  %   there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
  %   Since $\mathcal{M}$ is containment-free, we must have $b < a$.
  %   Then it follows that we have
  %   $\mu(b) < \mu(a)$ (since $\nu_1$ is increasing) and
  %   $\mu(b') > \mu(a')$ (since $\nu'_2$ is above $\nu'_2$),
  %   which contradicts our hypothesis about $\mathcal{M}$.
  %   \qed
  % \end{proof}
  %
  % \begin{claim}
  %   \label{claim:no (nu'_1, nu'_2)-edge}
  %   There is no $(\nu'_1, \nu'_2)$-edge in $\mathcal{M}$.
  % \end{claim}
  %
  % \begin{proof}[of Claim~\ref{claim:no (nu'_1, nu'_2)-edge}]
  %   Suppose, aiming at a contradiction that there is a
  %   $(\nu'_1, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
  %   According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
  %   there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
  %   Then it follows that we have
  %   $\mu(b) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
  %   $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu'_2$),
  %   which contradicts our hypothesis about $\mathcal{M}$.
  %   \qed
  % \end{proof}

  \begin{figure}[t!]
    \centering
      \begin{tikzpicture}
        [
          scale=0.35,
          label/.style={anchor=base}
        ]
        % nu_2
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,5.8) grid (5.2,11.2);
        \foreach \x/\y in {0/11,1/10,2/9,3/8,4/7,5/6} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % nu_3
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,15.8) grid (11.2,21.2);
        \foreach \x/\y in {6/16,7/17,8/18,9/19,10/20,11/21} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % nu'_2
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,-0.2) grid (17.2,5.2);

        \foreach \x/\y in {12/5,13/4,14/3,15/2,16/1,17/0} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % labels
        \node [label] (x) at (2,4.5) {$x$};
        \node [label] (a) at (8,14.5) {$a$};
        \node [label] (xp) at (14,-1.5) {$x'$};
        \node [label] (ap) at (16,-1.5) {$a'$};
        \node (nu1) at (2.5,2.5) {$\nu_2$};
        \node (nu2) at (8.5,12.5) {$\nu_3$};
        \node (nu1') at (14.5,-3.5) {$\nu'_2$};
        % crossings edges
        \draw [line width=1pt]
        (2,9) .. controls +(0,25) and +(0,30) .. (14,3);
        \draw [line width=1pt]
        (8,18) .. controls +(0,15) and +(0,20) .. (16,1);
      \end{tikzpicture}
    \caption{\label{fig:subfig:no (nu_3, nu'_2)-edge}%
      Claim~\ref{claim:no (nu_3, nu'_2)-edge}:
      Assuming a $(\nu_3, \nu'_2)$-edge $(a, b)$ in $\mathcal{M}$.
    }
  \end{figure}

  \begin{claim}
    \label{claim:no (nu_3, nu'_2)-edge}
    There is no $(\nu_3, \nu'_2)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no (nu_3, nu'_2)-edge}]
    Suppose, aiming at a contradiction, that there exists
    $(\nu_3, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
    According to Claim~\ref{claim:(nu_2, nu'_2)-edge}, there
    exists at least one $(\nu_2, \nu'_2)$-edge $(x, x')$ in $\mathcal{M}$.
    Since $\mathcal{M}$ is containment-free (Proposition~\ref{proposition:matching}),
    we have $x' < a'$.
    Therefore,
    we have $\mu(x) < \mu(a)$ (since $\nu_2$ is below $\nu_3$) and
    $\mu(x') > \mu(a')$ (since $\nu'_2$ is decreasing),
    which contradicts our hypothesis about $\mathcal{M}$.
    See Fig.~\ref{fig:subfig:no (nu_3, nu'_2)-edge}.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:at most one (nu_2, nu_3)-edge}
    There is at most one $(\nu_2, \nu_3)$-edge
    in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:at most one (nu_2, nu_3)-edge}]
    Apply Lemma~\ref{lemma:at most one edge monotone}.
    \qed
  \end{proof}

    We will see soon (upcoming Claim~\ref{claim:all (nu_2, nu'_2)-edges})
    that there exists actually no $(\nu_2, \nu_3)$-edge
    in $\mathcal{M}$.

  \begin{claim}
    \label{claim:no (nu_1, nu_3)-edge and no (nu'_1, nu_3)-edge}
    There is neither a $(\nu_1, \nu_3)$-edge not a $(\nu'_1, \nu_3)$-edge
    in $\mathcal{M}$.
  \end{claim}

 \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_3)-edge and no (nu'_1, nu_3)-edge}]
    We first prove that there is no $(\nu_1, \nu_3)$-edge in $\mathcal{M}$.
    Indeed, suppose, aiming at a contradiction, that these exists
    a $(\nu_1, \nu_3)$-edge $(a, a')$ in $\mathcal{M}$.
    According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
    there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
    Since $\mathcal{M}$ is containment-free, we must have
    $b < a'$.
    Then it follows that
    $\mu(b) < \mu(a)$ (since $\nu_1$ is incresing) and
    $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu_3$),
    which contradicts our hypothesis about $\mathcal{M}$.

    As for no $(\nu_1, \nu_3)$-edges in $\mathcal{M}$, suppose,
    aiming at a contradiction, that these exists
    a $(\nu'_1, \nu_3)$-edge $(a, a')$ in $\mathcal{M}$.
    According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
    there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
    Then it follows that
    $\mu(b) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
    $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu_3$),
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:(nu_3, nu'_3)-edge}
    There exists a $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:(nu_3, nu'_3)-edge}]
    Suppose, aiming at a contradiction, that there is no
    $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
    Combining
    Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge},
    Claim~\ref{claim:no (nu_3, nu'_2)-edge},
    Claim~\ref{claim:at most one (nu_2, nu_3)-edge}
    and Claim~\ref{claim:no (nu_1, nu_3)-edge and no (nu'_1, nu_3)-edge},
    we conclude that $N_3-1$ positions in $\nu_3$ are involved
    in edges of $\mathcal{M}$ that are
    neither ($\nu_1, \nu_3)$-edges,
    nor $(\nu_2, \nu_3)$-edges
    nor $(\nu'_1, \nu_3)$-edges
    nor $(\nu_3, \nu_3)$-edges.
    But $N_3 - 1 > 2N_4 + 2n + 2k + 4$ (the positions left) and hence
    $\mathcal{M}$ is not a perfect matching,
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:no edge below one (nu_3, nu'_3)-edge}
    There is
    neither a $(\sigma', \sigma')$-edge,
    nor a $(\sigma', \nu_4)$-edge,
    nor a $(\sigma', \nu'_2)$-edge,
    nor a $(\nu_4, \nu_4)$-edge
    nor a $(\nu_4, \nu'_2)$-edge,
    nor a $(\nu'_2, \nu'_2)$-edge
    in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no edge below one (nu_3, nu'_3)-edge}]
    Combine Claim~\ref{claim:(nu_3, nu'_3)-edge} with the fact that
    $\mathcal{M}$ is containment-free.
    \qed
  \end{proof}

  The next two claims state that $\mathcal{M}$ actually contains
  all $(\nu_1, \nu'_1)$-edges and all $(\nu_2, \nu'_2)$-edges.

  \begin{claim}
    \label{claim:all (nu_1, nu'_1)-edges}
    $\mathcal{M}$ contains $N_1$ pairwise crossing $(\nu_1, \nu'_1)$-edges.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:all (nu_1, nu'_1)-edges}]
    Suppose, aiming at a contradiction, that
    $\mathcal{M}$ does not contain $N_1$ $(\nu_1, \nu'_1)$-edge.
    Combing Claim~\ref{claim:no (nu_2, nu'_1)-edge} and
    Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge},
    we conclude that there exists a position $a$ in $\nu'_1$ that is
    involded in an edge $(a, a')$ in $\mathcal{M}$ that is
    neither a $(\nu_1, \nu'_1)$-edge,
    nor a $(\nu_2, \nu'_1)$-edge,
    nor a $(\nu'_1, \nu'_1)$-edge.
    (In other words, $a' > 2N_1 + N_2$.)
    Furthermore, according to Claim~\ref{claim:(nu_1, nu'_1)-edge},
    there exists a $(\nu_1, \nu'_1)$-edge $(x, x')$ in $\mathcal{M}$.
    Then, we have
    $\mu(x) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
    $\mu(x') < \mu(a')$ (since $\nu'_1$ is above any other gadget),
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:all (nu_2, nu'_2)-edges}
    $\mathcal{M}$ contains $N_2$ pairwise crossing $(\nu_2, \nu'_2)$-edges.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:all (nu_2, nu'_2)-edges}]
    The key idea is to focus on $\nu'_2$.
    Combining Claim~\ref{claim:no (nu_3, nu'_2)-edge},
    Claim~\ref{claim: no edge right of nu'_2} and
    Claim~\ref{claim:no edge below one (nu_3, nu'_3)-edge},
    we obtain that every position in $\nu'_2$ is involved in
    either a $(\nu_1, \nu'_2$)-edge
    or a $(\nu'_1, \nu'_2$)-edge
    or $(\nu_2, \nu'_2$)-edge.
    But, according to Claim~\ref{claim:all (nu_1, nu'_1)-edges},
    all positions in $\nu_1$ and $\nu'_1$ are involved in
    $(\nu_1, \nu'_1)$-edges.
    Then it follows that
    all positions in $\nu'_2$ are involved in a $(\nu_2, \nu'_2)$,
    and hence
    $\mathcal{M}$ contains $N_2$ pairwise crossing $(\nu_2, \nu'_2)$-edges.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:no (nu_2, nu_4)-edge}
    There is no $(\nu_2, \nu_4)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu_4)-edge}]
    According to Claim~\ref{claim: all (nu_2, nu'_2)-edges},
    all positions in $\nu_2$ and $\nu'_2$ are involved in
    $(\nu_2, \nu'_2)$-edges in $\mathcal{M}$.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:no (nu_4, nu'_3)-edge}
    There is no $(\nu_4, \nu'_3)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no (nu_4, nu'_3)-edge}]
    Suppose, aiming at a contradiction, that there exists
    a $(\nu_4, \nu'_3)$-edge $(a, a')$ in $\mathcal{M}$.
    According to Claim~\ref{claim:(nu_3, nu'_3)-edge},
    there exists a $(\nu_3, \nu'_3)$-edge $(x, x')$
    in $\mathcal{M}$.
    Since $\mathcal{M}$ is containment-free, we must have
    $x' < a'$.
    Therefore,
    we have $\mu(x) > \mu(a)$ (since $\nu_3$ is above $\nu_4$) and
    $\mu(x') < \nu(a')$ (since $\nu'_3$ is increasing),
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:(nu_4, nu'_4)-edge}
    There is at least one $(\nu_4, \nu'_4)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:(nu_4, nu'_4)-edge}]
    Suppose, aiming at a contradiction, that there is no
    $(\nu_4, \nu'_4)$-edge in $\mathcal{M}$.
    First, according to
    Claim~\ref{claim:all (nu_1, nu'_1)-edges} and
    Claim~\ref{claim:all (nu_2, nu'_2)-edges},
    there is neither a $(\nu_1, \nu_4)$-edge
    nor a $(\nu'_1, \nu_4)$-edge
    nor a $(\nu_2, \nu_4)$-edge
    nor a $(\nu_4, \nu'_2)$-edge in $\mathcal{M}$.
    Furthermore, according to
    Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge} and
    Claim~\ref{claim:no (nu_4, nu'_3)-edge},
    there is neither a $(\nu_3, \nu_4)$-edge
    nor a $(\nu_4, \nu'_3)$-edge in $\mathcal{M}$.
    But $N_4 > 2n + 2k + 2$, and
    hence $\mathcal{M}$ is not perfect,
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:no edge below (nu_4, nu'_4)-edge}
    There is
    neither a $(\pi', \pi')$-edge
    nor a $(\sigma', \pi'')$-edge
    nor a $(\sigma', \sigma'')$-edge
    in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge}]
    Combine Claim~\ref{claim:(nu_4, nu'_4)-edge} with the
    fact that $\mathcal{M}$ is containment-free.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:no (sigma', nu'_3)-edge, no (sigma', nu'_4)-edge}
    There is neither a $(\sigma', \nu'_3)$-edge
    nor a $(\sigma', \nu'_4)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\label{claim:no (sigma', nu'_3)-edge}]
    According to
    Claim~\ref{claim: no edge right of nu'_2},
    Claim~\ref{claim:all (nu_1, nu'_1)-edges}
    Claim~\ref{claim:all (nu_2, nu'_2)-edges} and
    Claim~\ref{claim:no (nu_4, nu'_3)-edge}
    every position in $\sigma'$ is involved in
    either a $(\sigma', \nu'_3)$-edge
    or a $(\sigma', \pi')$-edge
    or a $(\sigma', \nu'_4)$-edge
    in $\mathcal{M}$.
    Let $a$ be the first position of $\sigma'$
    (\emph{i.e.}, $a = 2N_1 + N_2 + N_3 + 1$)
    and
    $b$ be the last position of $\sigma'$
    (\emph{i.e.}, $b = 2N_1 + N_2 + N_3 + k + 2$),
    and write $(a, a')$ and $(b, b')$ for the two associated edges of $\mathcal{M}$.
    We consider the presence of a $(\sigma', \nu'_3)$-edge
    or a $(\sigma', \nu'_4)$-edge in $\mathcal{M}$ separately.
    \begin{itemize}
      \item
      Suppose, aiming at a contradiction, that there exists
      a $(\sigma', \nu'_3)$-edge in $\mathcal{M}$.
      Since $\mathcal{M}$ is containment-free, we conclude that
      $(a, a')$ must be a $(\sigma', \nu'_3)$-edge.
      But $\mu(a) = 2N_2 + N_4 + 2n + k + 2 + (k+1) <
      \mu(b) = 2N_2 + N_4 + 2n + k + 2 + (k+2)$, and hence
      $(b, b')$ is also a $(\sigma', \nu'_3)$-edge
      (otherwise we would have $\mu(a') > \mu(b')$ since
      $\nu'_3$ is above $\pi'$ and $\nu'_4$,
      which contradicts our hypothesis).
      Consider any position $c$ but $a$ and $b$ of $\sigma'$
      (\emph{i.e.},
      $a = 2N_1 + N_2 + N_3 + 1 < c < 2N_1 + N_2 + N_3 + k + 2 = b$),
      and let $(c, c')$ be the associate edge in $\mathcal{M}$.
      We first observe that $(c, c')$ is not a $(\sigma', \nu'_3)$-edge,
      since otherwise we would have
      $\mu(a) > \mu(c)$ and $\mu(a') < \mu(c')$
      (since $\mathcal{M}$ is containment-free and $\nu'_3$ is increasing),
      thereby contradicting our hypothesis about $\mathcal{M}$.
      Therefore, $(c, c')$ is either a $(\sigma', \pi)$-edge or a $(\sigma', \nu'_4)$-edge.
      In both cases, $(b, b')$ is contained in $(c, c')$,
      which contradicts our hypothesis about $\mathcal{M}$ being containment-free.
      \item
      Suppose, aiming at a contradiction, that there exists
      a $(\sigma', \nu'_4)$-edge in $\mathcal{M}$.
      Since $\mathcal{M}$ is containment-free, we conclude that
      $(b, b')$ must be a $(\sigma', \nu'_4)$-edge.
      Since there is no $(\sigma', \nu'_3)$-edge in $\mathcal{M}$
      (preceding point),
      $(a, a')$ is either a
      $(\sigma', \pi')$-edge or a $(\sigma', \nu'_4)$-edge.
      In both cases, we have $a' < a$.
      Then it follows that
      $\mu(a) < \mu(b)$
      (since $2N_2 + N_4 + 2n + k + 2 + (k+1) < 2N_2 + N_4 + 2n + k + 2 + (k+2)$)
      and
      $\mu(a') > \mu(b')$ (since $\pi'$ is above $\nu'_4$ if $(a, a')$ is
      a $(\sigma', \pi')$-edge and
      $\nu'_4$ is decreasing if $(a, a')$ is
      a $(\sigma', \nu'_4)$-edge),
      which contradicts our hypothesis about $\mathcal{M}$.
    \end{itemize}
    \qed
  \end{proof}

  Combining the above claims, we conclude that there are
  $k+2$ $(\sigma', \pi')$-edges in $\mathcal{M}$.
  Recall that
  $\sigma' = ((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2]$ and that
  $\pi' = ((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2]$.
  Then it follows we have
  at least $k$ (possibly $k+1$ or $k+2$) independent $(\sigma', \pi')$-edges
  $(a, a')$ in $\mathcal{M}$ with
  $2N_1 + N_2 + N_3 + 1 < a < 2N_1 + N_2 + N_3 + k + 2$
  and
  $2N_1 + N_2 + N_3 + (k + 2) + 1 < a' < 2N_1 + N_2 + N_3 + (k + 2) + n + 2$.
  Therefore, by our hypothesis about $\mathcal{M}$,
  $\sigma$ occurs as a pattern in $\pi$.
  \qed
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% %%
% %% ---- Patterns ----
% %%
% \section{Patterns}
% \label{section:Patterns}
%
% \begin{proposition}
%   \label{proposition:patterns:lower bound}
%   Let $\pi \in S_{n}$.
%   Then $\pi$ contains a pattern of size at least
%   $\left\lfloor\sqrt{n-1}/2\right\rfloor$ that
%   is the union of two order-isomorphic patterns.
% \end{proposition}
%
% \begin{proof}[of Proposition~\ref{proposition:patterns:lower bound}]
%   The Erdös-Szekeres theorem \cite{Erdos:Szekeres:1935}
%   states that every permutation of
%   length at least $pq+1$ must contain either the pattern
%   $\nearrow_{p+1}$ or the pattern $\searrow_{q+1}$.
%   The result now follows from the fact that
%   the pattern $\nearrow_{p+1}$ (resp. $\searrow_{q+1}$)
% \qed
% \end{proof}
%
% \begin{proposition}
%   \label{proposition:patterns:upper bound}
%   If $2^{\binom{k}{2}} > \binom{n}{2k}\;\binom{2k}{k}$, then there exists
%   a permutation of $S_{n}$
%   that does not contains a pattern
%   of length $2k$ that is the union of two
%   order-isomorphic patterns.
% \end{proposition}
%
% \begin{proof}[of Proposition~\ref{proposition:patterns:upper bound}]
% The proof is by the probabilistic method \cite{Alon:Spencer:1992}.
% Let $\pi \in S_n$ be chosen randomly and uniformly.
% For any pattern $\sigma$ of length $2k$ ($k$ to be precisely defined latter)
% of $\pi$,
% let $X_\sigma$ be the indicator variable that
% $\sigma$ is the union of two order-ismorphic patterns.
% According to Boole's inequality, we have
% $P\left(X_\sigma\right) \leq \binom{2k}{k}\;2^{-k}$.
% Since there are $\binom{n}{2k}$ patterns of length $2k$ in $\pi$,
% the probability that $\pi$ contains a pattern of length $2k$
% is the union of two order-ismorphic patterns is at most
% $\binom{n}{2k}\;\binom{2k}{k}\;2^{-k} < 1$.
% Thus, with positive probability, no event $X_\sigma$ occurs and
% there is a permutation of $S_n$ that does not contain a pattern
% of length $2k$ that is the union of two order-isomorphic patterns.
% \qed
% \end{proof}

\bigskip
\bigskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% ---- Conclusion ----
%%
\section{Conclusion}
\label{section:Conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Bibliography
%%

\bibliographystyle{plain}
\bibliography{biblio}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section*{Appendix (Reviewers' version only)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%